---
title: "Principle Component Aanalysis"
output: html_document
---
## Data
The data you will be using comes from the Assistments online intelligent tutoring system (https://www.assistments.org/). It describes students working through online math problems. Each student has the following data associated with them:

- id
- prior_prob_count: How many problems a student has answered in the system prior to this session
- prior_percent_correct: The percentage of problems a student has answered correctly prior to this session
- problems_attempted: The number of problems the student has attempted in the current session
- mean_correct: The average number of correct answers a student made on their first attempt at problems in the current session
- mean_hint: The average number of hints a student asked for in the current session
- mean_attempt: The average number of attempts a student took to answer a problem in the current session
- mean_confidence: The average confidence each student has in their ability to answer the problems in the current session

## Start by uploading the data
```{r}
D1 <- read.csv("Assistments-confidence.csv",header=TRUE)

  #We won't need the id variable, so remove that.

D1 <- D1[,-1]

```

## Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.

```{r}
#You can install the corrplot package to plot some pretty correlation matrices (sometimes called correlograms)
install.packages("corrplot")
library(corrplot)

#Generate pairwise correlations
COR <- cor(D1)

corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

#Study your correlogram image and save it, you will need it later
```

## Create a new data frame with the mean_correct variables removed

```{r}
D2 <- D1[,-4]

#The, scale and center your data for easier interpretation
D2 <- scale(D2, center = TRUE)
```

## Now run the PCA on the new data frame

```{r}
pca <- prcomp(D2, scale = TRUE)
```

## Although princomp does not generate the eigenvalues directly for us, we can print a list of the standard deviation of the variance accounted for by each component.

```{r}
pca$sdev

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue

pca$sdev^2

#A summary of our pca will give us the proportion of variance accounted for by each component

summary(pca)

#We can look at this to get an idea of which components we should keep and which we should drop

plot(pca, type = "lines")
```

## Think about which components you would drop and make a decision

# I think I will keep 3 principal components because the variance differences between the 4th, 5th and 6th components are relatively small. Or according to the rule of thumb, 5 principal components would explain over 90% of the variance. So 5 would also be an option.

## Part II

```{r}
#Now, create a data frame of the transformed data from your pca.

D3 <- as.data.frame(pca$x)

#Attach the variable "mean_correct" from your original data frame to D3.

D4 <- cbind(D3, as.data.frame(D1$mean_correct))

#Now re-run your scatterplots and correlations between the transformed data and mean_correct. If you had dropped some components would you have lost important infomation about mean_correct?

COR2 <- cor(D4)
corrplot(COR2, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

# I think it is ok if we drop PC3 and PC5. The information would not be lost too much.

```
## Now print out the eigenvectors (often called loadings) for the components you generated:

```{r}
pca$rotation

#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component

loadings <- abs(pca$rotation) #abs() will make all eigenvectors positive

sweep(loadings, 2, colSums(loadings), "/") #sweep() computes each row as a proportion of the column. (There must be a way to do this with dplyr()?)

#Now examine your components and try to come up with substantive descriptions of what some might represent?

# If we look at the loadings of prinipal to each variable and we want to keep 3 principal components, we may assume that:
# PC1 has higher loadings for problems_attempted, mean_hint and mean_attempt. This could indicate the students quick learning abilities or learning desire. Because this is highly correleated with their attempts to answer questions and getting hints to these questions.

# PC2 has higher loadings for prior_prob_count and prior_percent_correct. This could indicate the students prior knowledge to these questions.

# PC3 has higher loadings for prior_prob_count, problems_attempted and mean_confidence. This could indicate students' confidence or attitude towards studying. I believe this could be related to students characteristics or study methods.

#You can generate a biplot to help you, though these can be a bit confusing. They plot the transformed data by the first two components. Therefore, the axes represent the direction of maximum variance. Then mapped onto this point cloud are the original directions of the variables, depicted as red arrows. It is supposed to provide a visualization of which variables "go together". Variables that possibly represent the same underlying construct point in the same direction.  

biplot(pca)

#Calculate values for each student that represent these your composite variables and then create a new correlogram showing their relationship to mean_correct.

corrplot(COR2, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")



```
# Part III  
## Also in this repository is a data set and codebook from Rod Martin, Patricia Puhlik-Doris, Gwen Larsen, Jeanette Gray, Kelly Weir at the University of Western Ontario about people's sense of humor. Can you perform a PCA on this data?

```{r}
# Import data and manipulate

C1<-read.csv("humor_data.csv",header=TRUE)
C2<-C1[,1:32]

COR3 <- cor(C2)

corrplot(COR3, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

# It's a mess! Too many variables!

C3<-scale(C2,center=TRUE)

# Perform a PCA

pca3 <- prcomp(C3, scale = TRUE)
summary(pca3)
plot(pca3, type="lines")

# Since the elbow shows at the 6th components, we should choose 5 principal components to analyze these variables.

pca3$rotation

loadings3 <- abs(pca3$rotation) 

sweep(loadings3, 2, colSums(loadings),"/")

biplot(pca3)

C4<-as.data.frame(pca3$x)

C5<-cbind(C4[,1:5],C1[,33:36])
COR3<-cor(C5)

corrplot(COR3, method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")

loadings4<-loadings3[,1:5]
sweep(loadings4, 2, colSums(loadings),"/")

# Conclusions:
# Since there are too many questions, we will mainly analyze our results and principal components according to biplot and correlogram. In this study, I will choose 5 principal components to these questions.

# PC1 is correlated with affiliative, self-enhancing, aggressive and self-defeating. I think this would indicate the overall judgement of a person's humor, because it is correlated with most of the questions.

# PC2 would indicate those people who are self-defeating. Because the correlation between PC2 and self-defeating is high and positive.

# PC3 would indicate those people who are affiliative. Because the correlation between PC3 and affiliative is high and positive.

# PC4 would indicate those people who are self-enhancing. Because the correlation between PC4 and self-enhancing is high and positive.

# PC5 would indicate those people who are aggressive. Because the correlation between PC5 and aggressive is high and positive.

```

